The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-08 rational.tex
_

Now consider the following algebraic manipulation (assuming $a \neq 0$):
\[\begin{array}{rclll}
&& a & \mbox{initial~$a$} & (0) \\
&=& a * 1 & \mbox{$1$~is~identity~for~$*$} & (1) \\
&=& a * ((1/a) * a) & \mbox{$1/a$~is~the~multiplicative~inverse~of~$a$} & (2) \\
&=& (a * (1/a)) * a & \mbox{$*$~is~associative} & (3) \\
&=& 1 * a & \mbox{$1/a$~is~the~multiplicative~inverse~of~$a$} & (4) \\
&=& a & \mbox{$1$~is~identity~for~$*$} & (5)
\end{array}\]
As the semantics in Sec.~\ref{sec:rat} formalizes, the witness to this 
{{a <-> a}} isomorphism commputes as follows:
\begin{itemize}
\item We start at line (0) with the value {{20}};
\item We proceed to line (1) with the value {{(20,())}};
\item We proceed to line (2) with the value {{(20,((1/alpha),alpha))}} where
  $\alpha$ is essentially a logic variable;
\item We proceed to line (3) with the value {{((20,1/alpha),alpha)}};
\item At line (4) we require that $20$ and $1/\alpha$ cancel each other which
  forces the unification $\alpha=20$, i.e., the value becomes
  {{((20,1/20),20)}} which simplifies to {{((),20)}};
\item We reach line (5) with the value {{20}}.
\end{itemize}

@@

The general idea of the graphical notation is that combinators are
modeled by ``wiring diagrams'' or ``circuits''; values are modeled as
``particles'' that may appear on the wires; and evaluation is modeled
by the flow of particles along the wires.

>>>>

Now consider the following algebraic manipulation (assuming $a \neq 0$):
\[\begin{array}{rclll}
&& a & \mbox{initial~$a$} & (0) \\
&=& a * 1 & \mbox{$1$~is~identity~for~$*$} & (1) \\
&=& a * ((1/a) * a) & \mbox{$1/a$~is~the~multiplicative~inverse~of~$a$} & (2) \\
&=& (a * (1/a)) * a & \mbox{$*$~is~associative} & (3) \\
&=& 1 * a & \mbox{$1/a$~is~the~multiplicative~inverse~of~$a$} & (4) \\
&=& a & \mbox{$1$~is~identity~for~$*$} & (5)
\end{array}\]
As the semantics in Sec.~\ref{sec:rat} formalizes, the witness to this 
{{a <-> a}} isomorphism computes as follows:
\begin{itemize}
\item We start at line (0) with the value {{20}};
\item We proceed to line (1) with the value {{(20,())}};
\item We proceed to line (2) with the value {{(20,((1/alpha),alpha))}} where
  $\alpha$ is essentially a logic variable;
\item We proceed to line (3) with the value {{((20,1/alpha),alpha)}};
\item At line (4) we require that $20$ and $1/\alpha$ cancel each other which
  forces the unification $\alpha=20$, i.e., the value becomes
  {{((20,1/20),20)}} which simplifies to {{((),20)}};
\item We reach line (5) with the value {{20}}.
\end{itemize}

@@

The general idea of the graphical notation is that combinators are
modeled by ``wiring diagrams'' or ``circuits''; values are modeled as
``particles'' that may appear on the wires; and evaluation is modeled
by the flow of particles along the wires. In fact, when talking about
{{langRev}} circuits, we will often use the words {{value}} and
{{particle}] interchangeably.
